### 2024-11-26 paper/paper.qmd
update plots

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Xizi Sun
thanks: "Code and data are available at: https://github.com/LilianS77/US_Marriage."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(dplyr)
library(here)
library(arrow)
library(scales)
library(knitr)
library(broom)
library(rstanarm)
library(kableExtra)
library(broom.mixed)
```

```{r}
#| echo: false
#| warning: false
#| message: false

data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Data Tool

The dataset was analyzed using R [@citeR] and utilized several R packages for data manipulation and visualization, including ggplot2 [@citeGGplot2], dplyr [@citeDplyr], and here [@citeHere]. The data was processed using the Apache Arrow package [@citeArrow] for efficient file handling and visualized with scales [@citeScales] for enhanced graphics. Reproducibility was ensured using knitr [@citeKnitr]. Data was extracted from IPUMS USA [@ipumsusa]. Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].

## Data Source And Measurement

The data used for this study was sourced from **IPUMS USA (Integrated Public Use Microdata Series)** [@ipumsusa], a comprehensive repository providing access to harmonized census and survey data from the United States. IPUMS USA is renowned for its meticulous process of standardizing variables across datasets, ensuring compatibility over time, and providing extensive metadata to support research. This harmonization process enables researchers to conduct longitudinal and cross-sectional studies on social, demographic, and economic trends.

The dataset consists of **individual-level microdata**, where each record represents a single person, numerically coded for all relevant characteristics. These records are organized into households, allowing for the study of individual behaviors and characteristics in the context of their family or co-residential settings. Unlike compiled statistics or pre-aggregated tables, this microdata structure provides researchers with unparalleled flexibility in exploring relationships between variables.

To address the diversity of record layouts, coding schemes, and documentation across the historical scope of the dataset, IPUMS implements a rigorous **harmonization process**. Variables are assigned **uniform codes**, ensuring consistency across census years (1850–2010) and the American Community Surveys (ACS) (2000–present). This standardization simplifies the analysis of long-term trends and facilitates comparisons across time and space.

## Variable Selection

Using the IPUMS data extraction system, this study selected a focused subset of variables to examine the determinants of **non-marriage** in the United States. These include key demographic and socioeconomic characteristics, which are numerically coded for statistical analysis.

The table [@tbl-cleaned-data] shows the variables after data cleaning,

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned data
#| echo: false
#| warning: false
#| message: false

# Display the first 6 rows of the dataset with column names
head(data[, c("marital_status", "age", "gender", "Race", "Income", "education_level")], 6) |>
  kable(
    col.names = c("Marital Status", "Age", "Gender", "Race", "Income", "Education Level"),
    booktabs = TRUE
  )
```

### Outcome variables
The primary outcome variable for this study is **Marital Status**, which categorizes individuals based on their marital state. The proportion of marital status categories is displayed in figure @fig-proportion-marriage. This variable allows for a comparison between individuals who have never married (`Not_Married`) and those who have (`Married`). For this study,
**Not_Married**: Includes individuals who have never been married.
**Married**: Includes individuals who are married, as well as those who are divorced, widowed, or separated.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-proportion-marriage
#| fig-cap: Proportion of Marital Status Categories
#| 
# Calculate proportions
marital_status_counts <- table(data$marital_status)
marital_status_df <- as.data.frame(marital_status_counts)
colnames(marital_status_df) <- c("Marital_Status", "Count")
marital_status_df$Proportion <- marital_status_df$Count / sum(marital_status_df$Count)

# Plot
ggplot(marital_status_df, aes(x = Marital_Status, y = Proportion, fill = Marital_Status)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportion of Marital Status Categories",
    x = "Marital Status",
    y = "Proportion"
  ) +
  theme_minimal()
```

### Predictor variables

This section focuses on the predictor variables included in the study. The distribution of predictor variables is displayed in Figure @fig-predictor-variables. These variables capture demographic, socioeconomic, and personal characteristics, providing a comprehensive framework for analyzing factors associated with marital status. Below are the key predictor variables:

1. **Age**: A continuous variable representing the respondent's age in years.
2. **Gender**: A categorical variable indicating whether the respondent is male or female. Gender differences often play a role in marital patterns.
3. **Race**: A categorical variable categorized into White, Black, Asian, American Indian, and Other racial groups. This variable examines potential racial disparities in marital behavior.
4. **Education Level**: An ordinal variable indicating the highest level of education attained by the respondent. It is grouped into five categories: Below High School, High School, Some College, Bachelor’s Degree, and Above Bachelor.
5. **Income**: A continuous variable measuring the respondent's annual income in dollars. Income reflects economic resources and may be associated with marital stability and decisions.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: fig-predictor-variables
#| fig-subcap: ["Age Distribution", "Gender Distribution", "Race Distribution", "Education Level Distribution", "Income Distribution"]
#| label: fig-predictor-variables
#| layout-ncol: 2

# Function to create a bar plot or histogram
create_plot <- function(data, x_var, x_label, is_continuous = FALSE, binwidth = NULL) {
  if (is_continuous) {
    # Create a histogram for continuous variables
    p <- ggplot(data, aes_string(x = x_var)) +
      geom_histogram(binwidth = binwidth, fill = "skyblue", color = "black") +
      labs(x = x_label, y = "Count") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  } else {
    # Create a bar chart for categorical variables
    p <- ggplot(data, aes_string(x = x_var)) +
      geom_bar(fill = "skyblue", color = "black") +
      labs(x = x_label, y = "Count") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  }
  return(p)
}

# Create plots for the predictor variables
create_plot(data, "age", "Age", is_continuous = TRUE, binwidth = 5)
create_plot(data, "gender", "Gender")
create_plot(data, "Race", "Race")
create_plot(data, "education_level", "Education Level")
create_plot(data, "Income", "Income", is_continuous = TRUE, binwidth = 10000)

```

## Data Selection

To understand the phenomenon of non-marriage in the United States, I selected the IPUMS USA dataset over alternatives such as IPUMS International. While IPUMS International provides harmonized census data from 104 countries with over 1 billion person records, its vast scope makes it less targeted for this study. My focus is specifically on the U.S. population, and IPUMS USA offers high-precision data from American Community Surveys (ACS) and federal censuses, making it a more suitable choice for studying societal patterns specific to the United States.

# Model

For this analysis, I employed a Logistic Regression Model to assess the likelihood of an individual not being married (outcome variable) based on several demographic and socioeconomic predictors (predictor variables). This model was chosen due to the binary nature of the outcome variable, which distinguishes between individuals who are "Not Married" versus those who are married (including divorced, widowed, or separated).

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model Setup

### Objective
The primary objective of the model is to analyze and predict the factors associated with individuals' marital status, focusing on identifying key predictors for individuals who have never been married (`Not_Married`).

The logistic regression model used in this study is:

$$
\log \left( \frac{\hat{p}}{1 - \hat{p}} \right) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{gender} + \beta_3 \cdot \text{Race} + \beta_4 \cdot \text{Income} + \beta_5 \cdot \text{education\_level}
$$

The priors for the coefficients are defined as:

$$
\begin{aligned}
\beta_0 &\sim \text{Normal}(0, 2.5), \\
\beta_1 &\sim \text{Normal}(0, 2.5), \\
\beta_2 &\sim \text{Normal}(0, 2.5), \\
\beta_3 &\sim \text{Normal}(0, 2.5), \\
\beta_4 &\sim \text{Normal}(0, 2.5), \\
\beta_5 &\sim \text{Normal}(0, 2.5). \\
\end{aligned}
$$

Where:

- $\hat{p}$: Represents the probability that an individual is classified as `Not_Married`.
- $\beta_0$: The intercept term, representing the log-odds of being `Not_Married` when all predictors are zero.
- $\beta_1$: The coefficient for the individual's age.
- $\beta_2$: The coefficient for gender.
- $\beta_3$: The coefficient for the race category.
- $\beta_4$: The coefficient for annual income.
- $\beta_5$: The coefficient for the education level.

### Priors Explanation

In this model, normal priors with a mean of 0 and a standard deviation of 2.5 are assigned to all coefficients and the intercept. This choice reflects neutrality, implying no prior expectation of direction or magnitude for the coefficients.

The standard deviation of 2.5 reflects moderate uncertainty in prior beliefs, allowing the model to incorporate reasonable variability in predictor effects while avoiding overly restrictive priors.

## Model justification

The logistic regression model was chosen for this study due to its suitability for predicting binary outcomes, in this case, whether an individual is categorized as "Not Married" or not. Logistic regression is widely used in social science research to understand relationships between a binary dependent variable and multiple independent variables, making it an ideal choice for analyzing marital status. This model allows for a straightforward interpretation of the effects of predictors, such as age, income, education, and gender, through odds ratios, providing insight into the likelihood of being "Not Married." The dataset contains a mix of continuous and categorical variables, and logistic regression can handle this diversity effectively. Additionally, its theoretical simplicity ensures that the analysis remains interpretable while maintaining robustness. The choice of this model aligns with the study’s goal of identifying significant factors associated with non-marriage in the U.S., providing actionable insights while being grounded in established statistical methods. Despite its assumptions, such as linearity in the log-odds and independence of observations, logistic regression offers the flexibility to incorporate interaction terms and account for a range of predictor variables, making it a powerful tool for this research.

## Model Validation
From @tbl-model-coefficients, the validation indicates that the logistic regression model provides a robust understanding of the predictors of "Not Married" status. However, some predictors, such as race, show variability in their impact, necessitating cautious interpretation. The model's good fit, coupled with well-defined confidence intervals and validation metrics, ensures that it provides reliable insights into the determinants of marital status.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Load the saved final model
bayesian_model <-
  readRDS(file = here::here("models/bayesian_model.rds"))
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-model-coefficients
#| tbl-cap: Summary of the model

# Extract coefficients
coefficients <- broom::tidy(bayesian_model, conf.int = TRUE)
coefficients |> kable(digits = 2)
```

## Results

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-coefficient
#| fig-cap: Income Distribution for Unmarried Individuals

# Preprocess the data to create a binary 'Not_Married' variable
data$Not_Married <- ifelse(data$marital_status == "Not_Married", 1, 0)

# Fit the logistic regression model
model <- glm(Not_Married ~ age + gender + Race + Income + education_level, 
             data = data, family = binomial())

# Extract coefficients and confidence intervals
model_summary <- broom::tidy(model, conf.int = TRUE)

# Create the coefficient plot
ggplot(model_summary, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange(color = "orange") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() +
  labs(
    title = "Coefficients and Confidence Intervals",
    x = "Predictors",
    y = "Coefficient Value"
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Income-Unmarried
#| fig-cap: Income Distribution for Unmarried Individuals

# Filter for unmarried individuals
unmarried_data <- data %>% filter(marital_status == "Not_Married")

# Create the income distribution plot without scientific notation
ggplot(unmarried_data, aes(x = Income)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_density(color = "blue", size = 1) +
  ggtitle("Income Distribution for Unmarried Individuals") +
  xlab("Income") +
  ylab("Density") +
  scale_x_continuous(labels = comma) + # Format x-axis labels as regular numbers
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Age-Unmarried
#| fig-cap: Age Distribution for Unmarried Individua

# Filter for unmarried individuals
unmarried_data <- data %>% filter(marital_status == "Not_Married")

# Create the age distribution plot
ggplot(unmarried_data, aes(x = age)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "salmon", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +
  ggtitle("Age Distribution for Unmarried Individuals") +
  xlab("Age") +
  ylab("Density") +
  theme_minimal()
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-edu-marriage
#| fig-cap: Proportions of Marital Status by Education Level
#| 
# Calculate proportions of marital status within each education level
education_marital_status <- data %>%
  group_by(education_level, marital_status) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(education_level) %>%
  mutate(proportion = count / sum(count))

# Create the stacked bar chart with percentages and orange/blue colors
ggplot(education_marital_status, aes(x = education_level, y = proportion, fill = marital_status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = scales::percent(proportion, accuracy = 1)),
    position = position_stack(vjust = 0.5), size = 3.5, color = "white"
  ) +
  scale_fill_manual(values = c("Not_Married" = "#FFA500", "Married" = "#4682B4")) +
  labs(
    title = "Proportions of Marital Status by Education Level",
    x = "Education Level",
    y = "Proportion",
    fill = "Marital Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-race-marriage
#| fig-cap: Proportions of Marital Status by Race

# Calculate proportions of marital status within each race
race_marital_status <- data %>%
  group_by(Race, marital_status) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(Race) %>%
  mutate(proportion = count / sum(count))

# Create the stacked bar chart with annotations
ggplot(race_marital_status, aes(x = Race, y = proportion, fill = marital_status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    aes(label = scales::percent(proportion, accuracy = 1)),
    position = position_stack(vjust = 0.5), size = 3.5, color = "white"
  ) +
  scale_fill_manual(values = c("Not_Married" = "red", "Married" = "blue")) +
  labs(
    title = "Proportions of Marital Status by Race",
    x = "Race",
    y = "Proportion",
    fill = "Marital Status"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-gender-marriage
#| fig-cap: Marital Status Proportions by Gender

# Calculate proportions of marital status within each gender
gender_marital_status <- data %>%
  group_by(gender, marital_status) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  group_by(gender) %>%
  mutate(
    proportion = count / sum(count),
    percentage_label = paste0(scales::percent(proportion, accuracy = 1))
  )

# Separate pie charts for each gender with percentage labels
ggplot(gender_marital_status, aes(x = "", y = proportion, fill = marital_status)) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(
    aes(label = percentage_label),
    position = position_stack(vjust = 0.5), size = 4, color = "white"
  ) +
  coord_polar("y", start = 0) +
  facet_wrap(~gender) +
  scale_fill_manual(values = c("Not_Married" = "#FF9999", "Married" = "#66B3FF")) +
  labs(
    title = "Marital Status Proportions by Gender",
    x = NULL,
    y = NULL,
    fill = "Marital Status"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    strip.text = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```
# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix

## How to Extract 2022 ACS Data from IPUMS

To obtain data from IPUMS, we start by navigating to the IPUMS USA section and clicking on Get Data. Next, we go to the Select Sample section, where we uncheck the "Default sample from each year" option and instead select 2023 ACS. After selecting our sample, we proceed to add variables of interest. For individual-level data, we might add variables from the Person section. For example, under Demographic, we could include variables like AGE, and under Person, we could add SEX, RACE, INCTOT (total personal income) and EDUC (education attainment). Once our variables are selected, we click View Cart, then proceed by clicking Create Data Extract. At this point, we review our selections, change the Data Format to CSV, and submit our extract for processing. Then we saved it locally as usa_00001.csv.

## Model Assumptions
1. **Linearity**: The log odds of the outcome are linearly related to the predictors.
2. **Independence**: Observations are independent of each other.
3. **No Multicollinearity**: Predictors are not highly correlated.
4. **No Outliers or Influential Points**: Checked through residual analysis.

\newpage

# References